Northern Pikeminnow Predation on Juvenile Chinook Salmon and Potential Impacts to Local Populations

### 2021-11-12 analysis/paper/paper.Rmd
added some equations for several estimators

```{r schnabel-estimator, eval = T}
# using Schnabel model
N_mods = cmr_df %>%
  # filter out spring 2021 (no recapture)
  filter(event_name != "Spring_2021") %>%
  group_by(event_name, species, date) %>%
  summarise(n = n(),
            m = sum(capture_type == "Recapture", na.rm = T),
            u = sum(capture_type == "Mark" | is.na(capture_type)),
            R = sum(fish_status == "Live", na.rm = T),
            .groups = "drop") %>%
  group_by(event_name, species) %>%
  nest() %>%
  crossing(model = c("Schnabel",
                     "SchumacherEschmeyer")) %>%
  mutate(mr_model = map2(data,
                         model,
                         .f = function(x, y) {
                           with(x,
                                mrClosed(n = n,
                                         m = m,
                                         R = R,
                                         method = y,
                                         chapman.mod = TRUE))
                         }),
         N = map_dbl(mr_model,
                     .f = summary),
         CI = map(mr_model,
                  .f = confint),
         Lci = map_dbl(CI,
                       .f = function(x) x[1]), 
         Uci = map_dbl(CI,
                       .f = function(x) x[2])) %>%
  ungroup() %>%
  select(-CI)

@@

The Chapman-modified Lincoln-Peterson estimator is below, where $M$ is the number of fish marked and returned to the population, $n$ is the number of fish caught in the second/recapture event and $m$ is the number of marked fish in the second sample.
$$
\hat{N} = \frac{(M + 1)(n + 1)}{(m + 1)} - 1
$$
The Schnabel estimator is shown below, where the $M$, $n$ and $m$ are indexed by the sampling occasion, $i$. This estimator does not have an associate standard error, but 95% confidence intervals can be calculated.
$$
\hat{N} = \frac{\sum\limits_{i = 1}^k n_i M_i}{\left(\sum\limits_{i = 1}^k m_i \right) + 1}
$$

>>>>

```{r schnabel-estimator, eval = T}
# using Schnabel and Schumacher-Eschmeyer estimators
N_mods = cmr_df %>%
  # filter out spring 2021 (no recapture)
  filter(event_name != "Spring_2021") %>%
  group_by(event_name, species, date) %>%
  summarise(n = n(),
            m = sum(capture_type == "Recapture", na.rm = T),
            u = sum(capture_type == "Mark" | is.na(capture_type)),
            R = sum(fish_status == "Live", na.rm = T),
            .groups = "drop") %>%
  group_by(event_name, species) %>%
  nest() %>%
  crossing(model = c("Schnabel",
                     "SchumacherEschmeyer")) %>%
  mutate(mr_model = map2(data,
                         model,
                         .f = function(x, y) {
                           with(x,
                                mrClosed(n = n,
                                         m = m,
                                         R = R,
                                         method = y,
                                         chapman.mod = TRUE))
                         }),
         N = map_dbl(mr_model,
                     .f = summary),
         CI = map(mr_model,
                  .f = confint),
         Lci = map_dbl(CI,
                       .f = function(x) x[1]), 
         Uci = map_dbl(CI,
                       .f = function(x) x[2])) %>%
  ungroup() %>%
  select(-CI)

@@

The Lincoln-Petersen estimator is below, where $M$ is the number of fish marked and returned to the population, $n$ is the number of fish caught in the second/recapture event and $m$ is the number of marked fish in the second sample.

$$
\hat{N} = \frac{(M)(n)}{(m)}
$$

The Lincoln-Petersen estimator can be biased with small samples, so we also investigated the Chapman-modified Lincoln-Petersen estimator which is shown below.
$$
\hat{N} = \frac{(M + 1)(n + 1)}{(m + 1)} - 1
$$
The Schnabel estimator is shown below, where the $M$, $n$ and $m$ are indexed by the sampling occasion, $i$. In our example, the sampling occasions are defined as each day of sampling. This estimator does not have an associate standard error, but 95% confidence intervals can be calculated.
$$
\hat{N} = \frac{\sum\limits_{i = 1}^k n_i M_i}{\left(\sum\limits_{i = 1}^k m_i \right) + 1}
$$
There is another estimator for this type of multiple census surveys, called the Schumacher-Eschmeyer estimator, which is based on minimizing the weighted sum of squares between the proportion of marked individuals in the sample and the unknown proportion of marked individuals in the population. It is shown below.

$$
\hat{N} = \frac{\sum\limits_{i = 1}^k n_i M^2_i}{\sum\limits_{i = 1}^k m_i M_i}
$$
